---
title: "Introduction to arcpullr"
date: "`r format(Sys.time(), '%d %B, %Y')`"
vignette: >
  %\VignetteIndexEntry{Introduction to arcpullr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
output: 
  html_document:
    theme:
      version: 5
editor_options: 
  markdown: 
    wrap: 72
---

To see more complete package documentation check out:
<a href="https://pfrater.github.io/arcpullr/">
https://pfrater.github.io/arcpullr/</a>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
library(arcpullr)
```


## URL's for examples 
<a class="btn btn-primary" data-bs-toggle="collapse"href="#collapseExample" role="button" aria-expanded="false" aria-controls="collapseExample">
Example Source Data
</a>
<div class="collapse" id="collapseExample">
  <div class="card card-body"> 
```{r}
wdnr_base_url <- "https://dnrmaps.wi.gov/arcgis/rest/services"
streams_layer_url <- "WT_SWDV/WT_Inland_Water_Resources_WTM_Ext_v2/MapServer/2"
streams_url <- paste(wdnr_base_url, streams_layer_url, sep = "/")

# the mke_county polygon is available as an exported object in arcpullr
open_water_layer_url <- "WT_SWDV/WT_Inland_Water_Resources_WTM_Ext_v2/MapServer/3"
hydro_url <- paste(wdnr_base_url, open_water_layer_url, sep = "/")

wis_county_layer_url <- "https://datcpgis.wi.gov/arcgis/rest/services/Base/PolygonsExternal/MapServer/0"
```
</div>
</div>
<br>


# Feature (i.e. vector) Layers

## The `get_spatial_layer()` function

This is one of the core functions of the package, and allows users to pull Feature Service data from an ArcGIS REST API.

```{r, eval = FALSE}
# url <- some url from an ArcGIS REST API layer
layer <- get_spatial_layer(url)
```

The URL should be a specific layer from an ArcGIS REST API, such as the
[Wisconsin Dep't. of Natural Resources Musky Streams
layer](https://dnrmaps.wi.gov/arcgis/rest/services/WT_SWDV/WT_Designated_Waters_WTM_Ext/MapServer/19),
for example.

The `get_spatial_layer()` function will retrieve data from this layer,
and format it to an object of class \code{sf} (*i.e.* of the R package [sf:
Simple Features for R](https://r-spatial.github.io/sf/)).

This function is also query-able both using both SQL code and ArcGIS's
[Query (Feature
Service)](https://developers.arcgis.com/rest/services-reference/query-feature-service-.htm)
functionality.

### Querying Spatially via ArcGIS Feature Service

Along with SQL, layers from an ArcGIS REST API may be queried spatially.
This is accomplished with the `get_layer_by_*` family of functions.
These functions are essentially a wrapper around `get_spatial_layer`
that removes the abstraction of the spatial query syntax used by
ArcGIS. These functions require a spatial object of class `sf` to be
passed to the `geometry` argument. To test this out you can also quickly
create simple `sf` objects using `sf_lines()`, `sf_points()`, or
`sf_polygons()` to test out the service feature.

```{r, eval = FALSE}
mke_waters <- get_layer_by_poly(url = streams_url, mke_county)
```

Spatial queries can be done with polygons, lines, or points...just use
the respective `get_layer_by_*` function. See the 
[vignette on spatial querying](spatial_queries.html) for more detailed examples.

### Querying via SQL

To query via SQL within the function the field name for the query of
interest must be known. For example,

```{r, eval = FALSE}
wi_river <- get_spatial_layer(hydro_url, where = "WATERBODY_ROW_NAME = 'Wisconsin River'")
```

For multiple WHERE clauses see the `?sql_where` function, which automates the 
creation of a WHERE clause.

\
\

# Raster Layers

arcpullr is capable of pulling raster layers as well (denoted as map and image
layers by ArcGIS REST APIs). Utilize the `get_map_layer` and `get_image_layer`
functions to pull these respective layers. See more information in the 
[vignette on raster layers](raster_layers.html)

\
\

# Plotting Layers

Layers retrieved via any of the `get_*_layer` functions may be plotted using an 
appropriate method which corresponds to the object type. For quick plotting we've 
also included the `plot_layer` function. This function will plot layers using 
`ggplot2` by default or can be switched to base graphics using the `plot_pkg` 
argument. 

```{r pull_county_layer, warning = FALSE, echo = FALSE, eval = FALSE}
wis_counties <- get_spatial_layer(url = wis_county_url)
```

```{r show_layer_plotting, ref.block=c("pull_county_layer", "plot_county_layer"), eval = FALSE}
```

```{r plot_county_layer, fig.height = 7, fig.width = 7, echo = FALSE}
plot_layer(wis_counties)
```


